Estimating the association between Facebook adoption and well-being in 72 countries

### 2023-02-03 ms.Rmd
Edits

To better understand the plausible range of associations, we linked data tracking Facebook's global adoption with three indicators of well-being. We examined 72 countries' per capita active Facebook users in two age brackets (13-34 and 35+ years) as predictors of life satisfaction (LS), negative (NE), and positive psychological experiences (PE) at the level of years spanning 2008 to 2019. The well-being data represented 946,798 individuals' responses from the nationally representative Gallup World Poll Survey [@gallupHowDoesGallup2014]. We joined these unique datasets to conduct a descriptive study to answer three basic yet important questions: First, to what extent is Facebook adoption associated with well-being? Second, do these associations differ by age or sex. And finally, how might these associations have differed between countries? In addition, we were interested in whether the intensity of use might make a difference, and therefore conducted our analyses separately for daily active users and monthly active users. As a supplementary analysis, we replicated and present these analyses on meta-analytic mental health outcomes in an appendix.

@@

We studied two metrics of Facebook adoption, daily (DAU) and monthly active users (MAU), from 2008 to 2019 for 72 countries (see Figure \@ref(fig:fig-forest)). FB estimates yearly DAU & MAU with their corresponding means from June 1 to August 31, and rounds values greater than 10,000,000 to three significant digits, and values less than 10,000,000 to the nearest 10,000. Facebook calculates DAU and MAU estimates separately for individuals aged 13-34 and 35+. User age is determined based on Facebook profile information, which can be inaccurate (e.g. young users reporting an older age.) Accordingly, Facebook has trimmed 0.008% of total MAU to exclude accounts with unrealistic or non-reported ages.

@@

Facebook defines users as individuals who visited Facebook or used Messenger during the appropriate period. Although accurately captured, these numbers are not perfect indicators of actual user numbers because of possible duplicate and false accounts. Facebook estimates those accounts to account for 11% and 5% of global MAUs, respectively, and that the former may be more likely in developing regions. Because internal criteria and methodology for determining duplicate and false accounts can change over time, estimates of MAU and DAU can also change.

@@

Gallup World Poll (GWP) is a nationally representative annual survey of 1,000 civilian, non-institutionalised individuals aged 15 years or older from 164 countries conducted since 2005. The surveys are conducted face-to-face or on the phone in the respondents’ native language and by local interviewers (for details, see [@gallupHowDoesGallup2014]). We studied Positive (PE) and Negative Experiences (NE), and life satisfaction. PE and NE measure respondents’ experienced well-being on the day before the survey with five items each. For PE, these items are: "Did you feel well-rested yesterday?", "Were you treated with respect all day yesterday?", "Did you smile or laugh a lot yesterday?", "Did you learn or do something interesting yesterday?, and "(Did you experience the following feelings during a lot of the day yesterday?) How about enjoyment?".

The NE items are responses to "Did you experience the following feelings during a lot of the day yesterday?" for "physical pain", "worry", "sadness", "stress", and "anger". For analyses we used the means of these scales. Life satisfaction in the moment was measured with one 11-step Likert item, "Please imagine a ladder, with steps numbered from 0 at the bottom to 10 at the top. The top of the ladder represents the best possible life for you and the bottom of the ladder represents the worst possible life for you. On which step of the ladder would you say you personally feel you stand at this time?". For analyses, we scaled these variables to percentages, and aggregated the `r number(gwp_n, big.mark = ",")` individuals’ data to means and standard errors for each country, year, sex, and age (15-34 and 35+) combination (`r number(gwp_n_groups, big.mark = ",")` cells).

@@

Overall, per capita active Facebook users predicted positive experiences positively, and negative experiences negatively. These associations held when comparing countries to other countries as well as countries to themselves over time. Associations between Facebook adoption and life satisfaction were less certain within-countries, but strong when comparing countries to each other. While these descriptive results do not speak to causal effects, they align with other recent findings suggesting that technology use has not become increasingly associated with negative psychological outcomes over time [@vuorreThereNoEvidence2021], and that the increased adoption of internet technologies in general is not, overall, associated with widespread psychological harms [@vuorreGlobalWellbeingMental2022]. We also found that Facebook adoption predicted young demographics' positive well-being more strongly than it did older demographics', and that, although males' associations were generally slightly more positive, the sex differences in this dataset were very small and not credibly different from zero. These demography-based differences, and lack therein, were notable in light of previous literature that has reported young girls to be more at-risk of screen- and technology-based effects than young males [e.g. @kellySocialMediaUse2018; but see @kreskiSocialMediaUse2021].

@@

>>>>

To better understand the plausible range of associations, we conducted a descriptive study that linked data tracking Facebook's global adoption with three indicators of well-being. We examined 72 countries' per capita active Facebook users in two age brackets (13-34 and 35+ years) as predictors of life satisfaction (LS), negative (NE), and positive psychological experiences (PE) at the level of years spanning 2008 to 2019. The well-being data represented 946,798 individuals' responses from the nationally representative Gallup World Poll Survey [@gallupHowDoesGallup2014]. We joined these unique datasets to conduct a descriptive study to answer three basic yet important questions: First, to what extent is Facebook adoption associated with well-being? Second, do these associations differ by age or sex. And finally, how might these associations have differed between countries? In addition, we were interested in whether the intensity of use might make a difference, and therefore conducted our analyses separately for daily active users and monthly active users. As a supplementary analysis, we replicated and present these analyses on meta-analytic mental health outcomes in an appendix. Due to the exploratory and descriptive nature of our study, we did not have a priori hypotheses about the directions or magnitudes of the potential associations.

@@

We studied two metrics of Facebook adoption, daily (DAU) and monthly active users (MAU), from 2008 to 2019 for 72 countries (see Figure \@ref(fig:fig-forest)). FB estimates yearly DAU & MAU with their corresponding means from June 1 to August 31, and rounds values greater than 10,000,000 to three significant digits, and values less than 10,000,000 to the nearest 10,000. Facebook calculates DAU and MAU estimates separately for individuals aged 13-34 and 35+. User age is determined based on Facebook profile information, which can be inaccurate (e.g. young users reporting an older age.) Accordingly, Facebook has trimmed 0.008% of total MAU to exclude accounts with unrealistic or non-reported ages. FB chose the 13-34 and 35+ age categories in order to maximize the accuracy of the data. Nevertheless, the lower age category includes groups typically defined as adolescents and young adults [@sawyerAgeAdolescence2018].

@@

Facebook defines users as individuals who visited Facebook or used Messenger during the appropriate period. Any use (e.g. login) during the period was counted. Although accurately captured, these numbers are not perfect indicators of actual user numbers because of possible duplicate and false accounts. Facebook estimates those accounts to account for 11% and 5% of global MAUs, respectively, and that the former may be more likely in developing regions. Because internal criteria and methodology for determining duplicate and false accounts can change over time, estimates of MAU and DAU can also change.

@@

Gallup World Poll (GWP) is a nationally representative annual survey of 1,000 civilian, non-institutionalised individuals aged 15 years or older from 164 countries conducted since 2005. The surveys are conducted face-to-face or on the phone in the respondents’ native language and by local interviewers (for details, see [@gallupHowDoesGallup2014]). We studied Positive (PE) and Negative Experiences (NE), and life satisfaction. PE and NE measure respondents’ experienced well-being on the day before the survey with five items each. For PE, these items are: "Did you feel well-rested yesterday?", "Were you treated with respect all day yesterday?", "Did you smile or laugh a lot yesterday?", "Did you learn or do something interesting yesterday?, and "(Did you experience the following feelings during a lot of the day yesterday?) How about enjoyment?". The NE items are responses to "Did you experience the following feelings during a lot of the day yesterday?" for "physical pain", "worry", "sadness", "stress", and "anger". For analyses we used the means of these scales.

Prior studies using the NE and PE scales have found them to display acceptable validity and measurement invariance, and that response-style differences across world regions are small [@jebbSubjectiveWellBeingWorld2020]. In addition, Gallup claims that its questions are developed "using a global network of research and political scientists [The Brookings Institution, World Bank, USAID, United Nations, Daniel Kahneman, Ed Diener, Deepak Chopra, Richard Florida, John Helliwell, Jeffrey Sachs and Arthur Stone] who understand key issues concerning question development and construction and data gathering" [@gallupWorldPollMethodology2022, p.5]. Although not extensively validated, we believe these items and scales to be uniquely valuable for our goals due to their extensive scope across time, geo-, and demographics.

Life satisfaction in the moment was measured with one 11-step Likert item, the Cantril Self Anchoring Scale (or "Cantril ladder"): "Please imagine a ladder, with steps numbered from 0 at the bottom to 10 at the top. The top of the ladder represents the best possible life for you and the bottom of the ladder represents the worst possible life for you. On which step of the ladder would you say you personally feel you stand at this time?" [@cantril1965pattern]. In addition to inclusion in the data, we were motivated to use this scale due to its widespread use [@oecdHowLife20202020], and because of prior work establishing its reliability and validity [@kapteynDimensionsSubjectiveWellbeing2015].

For analyses, we scaled these variables to percentages, and aggregated the `r number(gwp_n, big.mark = ",")` individuals’ data to means and standard errors for each country, year, sex, and age (15-34 and 35+) combination (`r number(gwp_n_groups, big.mark = ",")` cells).

@@

Overall, per capita active Facebook users predicted positive experiences positively, and negative experiences negatively. These associations held when comparing countries to other countries as well as countries to themselves over time. Associations between Facebook adoption and life satisfaction were less certain within-countries, but strong when comparing countries to each other. While these descriptive results do not speak to causal effects, they align with other recent findings suggesting that technology use has not become increasingly associated with negative psychological outcomes over time [@vuorreThereNoEvidence2021], and that the increased adoption of internet technologies in general is not, overall, associated with widespread psychological harms [@vuorreGlobalWellbeingMental2022]. We also found that Facebook adoption predicted young demographics' positive well-being more strongly than it did older demographics', and that, although males' associations were generally slightly more positive, the sex differences in this dataset were very small and not credibly different from zero. These demography-based differences, and lack therein, were notable in light of previous literature that has reported young girls to be more at-risk of screen- and technology-based effects than young males [e.g. @kellySocialMediaUse2018; but see @kreskiSocialMediaUse2021]. However, those studies focused on slightly younger individuals (from 10 to 15 years old), which could at least partly explain the different findings.

@@

We also highlight the fact that while Facebook adoption remains the overall dominant social media platform, our results do not necessarily generalize across different platforms. For instance, in the United States, 13 to 17 year olds are more likely to use TikTok, Instagram, and Snapchat than Facebook, so the user base of FB now consists of relatively more older individuals [@pewresearchcenterTeensSocialMedia2022]. In addition to the demographic shifts between social media platforms, the platforms themselves change over time and their associations with well-being might therefore not be consistent over time or different countries [@vuorreThereNoEvidence2021]. Moreover, studying a single platform cannot provide a complete picture of the overall associations that social media as a whole might have with well-being, because different platforms are used by different people and for different purposes, all of which might serve to moderate any potential associations.
